Arbitrary Precision and Arbitrary $n$-grams

### 2011-02-28 paper/paper.Rnw
plots in paper
more paper writing

\begin{document}
\maketitle
\section*{Sentence Generation}

@@

		Large $n$-grams (particularly $n > 4$) tend to result in whole
		Shakespeare sentence fragments.  For example, ``O , my offence is rank ,
		it smells to heaven ; It hath the primal eldest curse upon't , A
		brother's murther !'' was generated by a hexagram, and is a direct quote
		from Hamlet, act \osn{3} scene \osn{3}.  This is a well-documented
		phenomenon, and is due to the fact that it's quite possible that the
		preceding $n-1$-gram appears only once in the entire corpus, so the
		sentence generator has no choice but to choose the next word in the
		quotation.

@@

>>>>

<<echo=F>>=
source('analysis.R')
@

\begin{document}

\maketitle
\section*{Implementation}

	Our design is built around dictionaries for each n-gram we consider.  We
	break the input text into words and add sentence separators (although we do
	not break the sentences apart, so trigrams or higher can cross sentence
	boundaries), and then process the list of words, tallying the number of each
	$n$-gram seen in the dictionary.

	We also can optionally account for unknown words by replacing the first
	instance of each word with an unknown symbol; this is simply an additional
	processing step applied to the raw list of words.

	We compute all $n$-gram count dictionaries up to the desired $n$-gram
	length, and use these to compute the probabilities of each $n$-gram.  This
	dictionary of probabilities is used to construct sentences and to measure
	perplexity.  We either use logarithmic or arbitrary-precision arithmetic
	depending on what the user specifies.

	We implemented Good-Turing smoothing, which adjusts the $n$-gram counts
	before they are turned into probabilities.  Like unknown-handling, this is
	simply a pipeline processing step.

	Sentence generation begins with a beginning of sentence symbol, and then
	chooses words from the $n$-gram dictionary corresponding to the minimum of
	the desired $n$-gram length or the number of words we have already created.

	Perplexity works similarly, scanning the input list and computing the
	running sum or product probability, depending on whether we are using
	arbitrary precision or logarithmic arithmetic.  Because computing the
	perplexity for each word is simply a matter of looking up a tuple in the
	appropriate dictionary (possibly containing an unknown), this step is fairly
	rapid.  If we do not have the appropriate $n$-gram in the dictionary, we
	first try replacing the last word with an unknown, and, if that fails, we
	shorten the list of words until a match is found, which, at shortest, is a
	single unknown by itself.

	\\

	The dictionary approach was simple and intuitive, but fairly slow; the
	generation of sentences spends the majority of its time on dictionary
	operations, and uses only moderate amounts of memory.  Much of the time in
	sentence generation is spent computing the list of $n$-gram options to
	choose from, since this requires, in our implementation, scanning through
	the entire $n$-gram dictionary to find matches for the $(n-1)$-gram.  If we
	precomputed this as we populated the dictionary, substantial time could be
	saved, although choosing a random element from that list still takes $O(n)$
	time.

\section*{Sentence Generation}

@@

		Large $n$-grams (particularly $n > 4$) tend to result in whole
		Shakespeare sentence fragments.  For example, ``O , my offence is rank ,
		it smells to heaven ; It hath the primal eldest curse upon't , A
		brother's murther !'' was generated by a hexagram, and is a direct quote
		from Hamlet, act \osn{3} scene \osn{3}.  This is a well-documented
		phenomenon, and is due to the fact that it's quite possible that the
		preceding $(n-1)$-gram appears only once in the entire corpus, so the
		sentence generator has no choice but to choose the next word in the
		quotation.

@@

\begin{figure}
\begin{center}
<<echo=F,fig=T>>=
pp_gen.plot(results.tolstoy)
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
<<echo=F,fig=T>>=
pp_gen.plot(results.shakespeare)
@
\end{center}
\end{figure}
